Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi

### 2013-10-07 writeup/hat-analysis.tex
Final edits

HAT systems provide transactions with transactional availability and
either high availability or sticky availability. They offer latency
and availability benefits over traditional distributed databases, yet
they cannot achieve all possible semantics. In this section, we
delineate ACID, distributed replica consistency, and session
consistency levels which can be achieved with high availability (Read
Committed isolation, variants of Repeatable Read, atomic reads, and
many session guarantees), those with sticky availability (read your
writes, PRAM and causal consistency), and properties that cannot be
provided in a HAT system (those preventing Lost Update and Write Skew
or guaranteeing recency).  We present a full summary of these results
in Section~\ref{sec:hat-summary}.

@@

\vspace{.5em}\noindent\textbf{{Monotonic writes}} requires that each
session's writes become visible in the order they were submitted. Any
order on transactions should also be consistent with any precedence
that a global observer would see.

\vspace{.5em}\noindent\textbf{{Writes Follow Reads}} requires that, if
a session observes an effect of transaction $T_1$ and subsequently
commits transaction $T_2$, then another session can only observe
effects of $T_2$ if it can also observe $T_1$'s effects (or later
values that supersede $T_1$'s); this corresponds to Lamport's
``happens-before'' relation~\cite{lamportclocks}.  Any order on
transactions should respect the transitive reads-from
order.\vspace{.5em}

@@

We illustrate the hierarchy of available, sticky available, and
unavailable consistency models we have discussed in
Figure~\ref{fig:hatcompared}. Many models are simultaneously
achievable, but we find several particularly compelling. If we combine
all HAT and sticky guarantees, we have transactional, causal snapshot
reads (i.e., Causal Transactional Predicate Cut Isolation). If we
combine MAV and P-CI, we have transactional snapshot reads. We can
achieve RC, MR, and RYW by simply sticking clients to servers. We can
also combine unavailable models---for example, an unavailable system
might provide PRAM and One-Copy
Serializability~\cite{daudjee-session}.

To the best of our knowledge, this is the first unification of
database ACID transactions, distributed consistency, and session
guarantee models. Interestingly, strong one-copy serializability
entails all other models, while considering the (large) power set of
all compatible models (e.g., the diagram depicts 144 possible HAT
combinations) hints at the vast expanse of consistency models found in
the literature. This taxonomy is not exhaustive
(Section~\ref{sec:conclusion}), but we believe it lends substantial
clarity to the relationships between a large subset of the prominent
ACID and distributed consistency models. Additional read/write
transaction semantics that we have omitted should be classifiable
based on the available primitives and HAT-incompatible anomaly
prevention we have already discussed.

In light the of current practice of deploying weak isolation levels
(Section~\ref{sec:modernacid}), it is perhaps surprising that so many
weak isolation levels are achievable as HATs. Indeed, isolation levels
such as Read Committed expose and are defined in terms of end-user
anomalies that could not arise during serializable execution. However,
the prevalence of these models suggests that, in many cases,
applications can tolerate these their associated anomalies. Given our
HAT-compliance results, this in turn hints that--despite
idiosyncrasies relating to concurrent updates and data recency--highly
available database systems can provide sufficiently strong semantics
for many applications. Indeed, HAT databases may expose more anomalies
than a single-site database operating under weak isolation
(particularly during network partitions). However, for a fixed
isolation level (which, in practice, can vary across databases and may
differ from implementation-agnostic definitions in the literature),
users of single-site database are subject to the same (worst-case)
application-level anomalies as a HAT implementation. The necessary
(indefinite) visibility penalties (i.e., the right side of
Figure~\ref{fig:hatcompared}) and lack of support for preventing
concurrent updates (via the upper left half of
Figure~\ref{fig:hatcompared}) mean HATs are \textit{not} well-suited
for all applications (see Section~\ref{sec:evaluation}): these
limitations are fundamental. However, common practices such as ad-hoc,
user-level compensation and per-statement isolation ``upgrades''
(e.g., \texttt{SELECT FOR UPDATE} under weak isolation)--commonly used
to augment weak isolation--are also applicable in HAT systems
(although they may in turn compromise availability).

>>>>

HAT systems provide transactions with transactional availability or
sticky transactional availability. They offer latency and availability
benefits over traditional distributed databases, yet they cannot
achieve all possible semantics. In this section, we describe ACID,
distributed replica consistency, and session consistency levels which
can be achieved with high availability (Read Committed isolation,
variants of Repeatable Read, atomic reads, and many session
guarantees), those with sticky availability (read your writes, PRAM
and causal consistency). We also discuss properties that cannot be
provided in a HAT system (those preventing Lost Update and Write Skew
or guaranteeing recency).  We present a full summary of these results
in Section~\ref{sec:hat-summary}.

@@

\vspace{.5em}\noindent\textbf{{Monotonic writes}} requires that each
session's writes become visible in the order they were submitted. Any
order on transactions (as in Read Uncommitted isolation) should also
be consistent with any precedence that a global observer would see.

\vspace{.5em}\noindent\textbf{{Writes Follow Reads}} requires that, if
a session observes an effect of transaction $T_1$ and subsequently
commits transaction $T_2$, then another session can only observe
effects of $T_2$ if it can also observe $T_1$'s effects (or later
values that supersede $T_1$'s); this corresponds to Lamport's
``happens-before'' relation~\cite{lamportclocks}.  Any order on
transactions should respect this transitive order.\vspace{.5em}

@@

We illustrate the hierarchy of available, sticky available, and
unavailable consistency models we have discussed in
Figure~\ref{fig:hatcompared}. Many models are simultaneously
achievable, but we find several particularly compelling. If we combine
all HAT and sticky guarantees, we have transactional, causally
consistent snapshot reads (i.e., Causal Transactional Predicate Cut
Isolation). If we combine MAV and P-CI, we have transactional snapshot
reads. We can achieve RC, MR, and RYW by simply sticking clients to
servers. We can also combine unavailable models---for example, an
unavailable system might provide PRAM and One-Copy
Serializability~\cite{daudjee-session}.

To the best of our knowledge, this is the first unification of
transactional isolation, distributed consistency, and session
guarantee models. Interestingly, strong one-copy serializability
entails all other models, while considering the (large) power set of
all compatible models (e.g., the diagram depicts 144 possible HAT
combinations) hints at the vast expanse of consistency models found in
the literature. This taxonomy is not exhaustive
(Section~\ref{sec:conclusion}), but we believe it lends substantial
clarity to the relationships between a large subset of the prominent
ACID and distributed consistency models. Additional read/write
transaction semantics that we have omitted should be classifiable
based on the available primitives and HAT-incompatible anomaly
prevention we have already discussed.

In light the of current practice of deploying weak isolation levels
(Section~\ref{sec:modernacid}), it is perhaps surprising that so many
weak isolation levels are achievable as HATs. Indeed, isolation levels
such as Read Committed expose and are defined in terms of end-user
anomalies that could not arise during serializable execution. However,
the prevalence of these models suggests that, in many cases,
applications can tolerate these their associated anomalies. Given our
HAT-compliance results, this in turn hints that--despite
idiosyncrasies relating to concurrent updates and data recency--highly
available database systems can provide sufficiently strong semantics
for many applications. Indeed, HAT databases may expose more anomalies
than a single-site database operating under weak isolation
(particularly during network partitions). However, for a fixed
isolation level (which, in practice, can vary across databases and may
differ from implementation-agnostic definitions in the literature),
users of single-site database are subject to the same (worst-case)
application-level anomalies as a HAT implementation. The necessary
(indefinite) visibility penalties (i.e., the right side of
Figure~\ref{fig:hatcompared}) and lack of support for preventing
concurrent updates (via the upper left half of
Figure~\ref{fig:hatcompared}) mean HATs are \textit{not} well-suited
for all applications (see Section~\ref{sec:evaluation}): these
limitations are fundamental. However, common practices such as ad-hoc,
user-level compensation and per-statement isolation ``upgrades''
(e.g., \texttt{SELECT FOR UPDATE} under weak isolation)---commonly used
to augment weak isolation---are also applicable in HAT systems
(although they may in turn compromise availability).

### 2013-10-07 writeup/hat-vldb.tex
more writing
Final edits

\newif\iftechreport
%\techreporttrue
\techreportfalse

@@

@@

\title{Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi}
{\author{Peter Bailis, Aaron Davidson, Alan Fekete{\fontsize{12}{14}$^\dagger$}, Ali Ghodsi, Joseph M. Hellerstein, Ion Stoica \\{\affaddr{UC Berkeley and {\fontsize{12}{14}$^\dagger$}University of Sydney}}}}
\maketitle

@@

\vspace{.5em}\noindent\textbf{Acknowledgments} We would like to thank
Peter Alvaro, Neil Conway, Evan Jones, Adam Oliner, Aurojit Panda,
Shivaram Venkataraman, and the HotOS and VLDB reviewers for their
helpful feedback on this work. This research was supported in part by
the Air Force Office of Scientiﬁc Research (grant FA95500810352),
DARPA XData Award FA8750-12-2-0331, the National Science Foundation
(grants CNS-0722077, IIS-0713661, and IIS-0803690), NSF CISE
Expeditions award CCF-1139158, the National Science Foundation
Graduate Research Fellowship (grant DGE-1106400), and by gifts from
Amazon Web Services, Google, SAP, Cisco, Clearstory Data, Cloudera,
Ericsson, Facebook, FitWave, General Electric, Hortonworks, Huawei,
Intel, Microsoft, NetApp, Oracle, Samsung, Splunk, VMware, WANdisco,
and Yahoo!.

>>>>

\newif\iftechreport
\techreporttrue
%\techreportfalse

@@

\pagestyle{empty}

@@

\title{Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi}
{\author{Peter Bailis, Aaron Davidson, Alan Fekete{\fontsize{12}{14}$^\dagger$}, Ali Ghodsi, Joseph M. Hellerstein, Ion Stoica \\[1mm]{\affaddr{UC Berkeley and {\fontsize{12}{14}$^\dagger$}University of Sydney}}}}
\maketitle

@@

\vspace{.5em}\noindent\textbf{Acknowledgments} We would like to thank
Peter Alvaro, Neil Conway, Evan Jones, Adam Oliner, Aurojit Panda,
Shivaram Venkataraman, and the HotOS and VLDB reviewers for their
helpful feedback on this work. This research was supported in part by
the Air Force Office of Scientiﬁc Research (grant FA95500810352),
DARPA XData Award FA8750-12-2-0331, the National Science Foundation
(grants CNS-0722077, IIS-0713661, and IIS-0803690), NSF CISE
Expeditions award CCF-1139158, the National Science Foundation
Graduate Research Fellowship (grant DGE-1106400), and by gifts from
Amazon Web Services, Google, SAP, Cisco, Clearstory Data, Cloudera,
Ericsson, Facebook, FitWave, General Electric, Hortonworks, Huawei,
Intel, Microsoft, NetApp, Oracle, Samsung, Splunk, VMware, WANdisco,
and Yahoo!.\vspace{.5em}

### 2013-10-07 writeup/intro.tex
Final edits

The last decade has seen a shift in the design of popular large-scale
database systems, from the use of transactional
RDBMSs~\cite{bernstein-book, gray-isolation, gray-virtues} to the
widespread adoption of loosely consistent distributed key-value
stores~\cite{bigtable, pnuts, dynamo}. Core to this shift was the 2000
introduction of Brewer's CAP Theorem, which stated that a highly
available system cannot provide ``strong'' consistency guarantees in
the presence of network partitions~\cite{brewer-slides}. As formally
proven~\cite{gilbert-cap}, the CAP Theorem pertains to a data
consistency model called linearizability, or the ability to read the
most recent write to a data item that is replicated across
servers~\cite{herlihy-art}. However, despite its narrow scope, the CAP
Theorem is often misconstrued as a broad result regarding the ability
to provide ACID database properties with high
availability~\cite{hat-hotos,brewer-slides, foundation-article}; this
misunderstanding has led to substantial confusion regarding replica
consistency, transactional isolation, and high availability. The
recent resurgence of transactional systems suggests that programmers
value transactional semantics, but, unfortunately, existing
transactional data stores do not provide availability in the presence
of partitions~\cite{orleans,foundation-article, hstore,spanner,eiger,
  walter,calvin}.

Indeed, serializable transactions---the gold standard of traditional
ACID databases---are not achievable with high availability in the
presence of partitions~\cite{davidson-survey}. However, database
systems have a long tradition of providing weaker isolation and
consistency guarantees~\cite{adya, ansicritique, gray-virtues,
  gray-isolation, kemme-thesis}. Today's ACID and NewSQL databases
often employ weak isolation models due to concurrency and performance
benefits; weak isolation is overwhelmingly the default setting in
these stores and is often the only option offered
(Section~\ref{sec:modernacid}). While weak isolation levels do not
provide serializability for general-purpose transactions, they are
apparently strong enough to deliver acceptable behavior to many
application programmers and are substantially stronger than the
semantics provided by current highly available systems. This raises a
natural question: which semantics can be provided with high
availability?

@@

Our main contributions in this paper are as follows. We relate the
many previously proposed isolation and consistency models to the goal
of high availability, which guarantees a response from each non-failed
data replica in the presence of arbitrary partitions between them.  We
classify which among the wide array of models are achievable with high
availability, denoting them as {\em Highly Available Transactions}
(HATs). In doing so, we demonstrate that although many implementations
of HAT semantics are not highly available, this is an artifact of the
implementations rather than an inherent property of the semantics. Our
investigation shows that, besides serializability, Snapshot Isolation
and Repeatable Read isolation are not HAT-compliant, while most other
isolation levels are achievable with high availability. We also
demonstrate that many weak replica consistency models from distributed
systems are both HAT-compliant and simultaneously achievable with
several ACID properties.

@@

At a high level, the virtues of HATs are guaranteed responses from any
replica, low latency, and a range of semantic guarantees including
several whose usefulness is widely accepted such as Read
Committed. However, highly available systems are fundamentally unable
to prevent concurrent updates to shared data items and cannot provide
recency guarantees for reads. To understand when these virtues and
limitations are relevant in practice, we survey both practitioner
accounts and academic literature, perform experimental analysis on
modern cloud infrastructure, and analyze representative applications
for their semantic requirements. Our experiences with a HAT prototype
running across multiple geo-replicated datacenters indicate that HATs
offer a one to three order of magnitude latency decrease compared to
traditional distributed serializability protocols, and they can
provide acceptable semantics for a wide range of programs, especially
those with monotonic logic and commutative updates~\cite{calm,
  crdt}. HAT systems can also enforce arbitrary foreign key
constraints for multi-item updates and, in some cases, provide limited
uniqueness guarantees. However, HATs can fall short for applications
with concurrency-sensitive operations, requiring non-HA coordination.

@@

\item For correct behavior, applications may require a combination of
  HAT and (ideally sparing use of) non-HAT isolation levels; future
  database designers should act accordingly.
\end{introenumerate}

>>>>

The last decade has seen a shift in the design of popular large-scale
database systems, from the use of transactional
RDBMSs~\cite{bernstein-book, gray-isolation, gray-virtues} to the
widespread adoption of loosely consistent distributed key-value
stores~\cite{bigtable, pnuts, dynamo}. Core to this shift was the 2000
introduction of Brewer's CAP Theorem, which stated that a highly
available system cannot provide ``strong'' consistency guarantees in
the presence of network partitions~\cite{brewer-slides}. As formally
proven~\cite{gilbert-cap}, the CAP Theorem pertains to a data
consistency model called linearizability, or the ability to read the
most recent write to a data item that is replicated across
servers~\cite{herlihy-art}. However, despite its narrow scope, the CAP
Theorem is often misconstrued as a broad result regarding the ability
to provide ACID database properties with high
availability~\cite{hat-hotos,brewer-slides, foundation-article}; this
misunderstanding has led to substantial confusion regarding replica
consistency, transactional isolation, and high availability. The
recent resurgence of transactional systems suggests that programmers
value transactional semantics, but most existing transactional data
stores do not provide availability in the presence of
partitions~\cite{orleans,foundation-article, hstore,spanner,eiger,
  walter,calvin}.

Indeed, serializable transactions---the gold standard of traditional
ACID databases---are not achievable with high availability in the
presence of network partitions~\cite{davidson-survey}. However, database
systems have a long tradition of providing weaker isolation and
consistency guarantees~\cite{adya, ansicritique, gray-virtues,
  gray-isolation, kemme-thesis}. Today's ACID and NewSQL databases
often employ weak isolation models due to concurrency and performance
benefits; weak isolation is overwhelmingly the default setting in
these stores and is often the only option offered
(Section~\ref{sec:modernacid}). While weak isolation levels do not
provide serializability for general-purpose transactions, they are
apparently strong enough to deliver acceptable behavior to many
application programmers and are substantially stronger than the
semantics provided by current highly available systems. This raises a
natural question: which semantics can be provided with high
availability?

@@

Our main contributions in this paper are as follows. We relate the
many previously proposed database isolation and data consistency
models to the goal of high availability, which guarantees a response
from each non-failing server in the presence of arbitrary network
partitions between them.  We classify which among the wide array of
models are achievable with high availability, denoting them as {\em
  Highly Available Transactions} (HATs). In doing so, we demonstrate
that although many implementations of HAT semantics are not highly
available, this is an artifact of the implementations rather than an
inherent property of the semantics. Our investigation shows that,
besides serializability, Snapshot Isolation and Repeatable Read
isolation are not HAT-compliant, while most other isolation levels are
achievable with high availability. We also demonstrate that many weak
replica consistency models from distributed systems are both
HAT-compliant and simultaneously achievable with several ACID
properties.

@@

At a high level, the virtues of HATs are guaranteed responses from any
replica, low latency, and a range of semantic guarantees including
several whose usefulness is widely accepted such as Read
Committed. However, highly available systems are fundamentally unable
to prevent concurrent updates to shared data items and cannot provide
recency guarantees for reads. To understand when these virtues and
limitations are relevant in practice, we survey both practitioner
accounts and academic literature, perform experimental analysis on
modern cloud infrastructure, and analyze representative applications
for their semantic requirements. Our experiences with a HAT prototype
running across multiple geo-replicated datacenters indicate that HATs
offer a one to three order of magnitude latency decrease compared to
traditional distributed serializability protocols, and they can
provide acceptable semantics for a wide range of programs, especially
those with monotonic logic and commutative updates~\cite{calm,
  crdt}. HAT systems can also enforce arbitrary foreign key
constraints for multi-item updates and, in some cases, provide limited
uniqueness guarantees. However, HATs can fall short for applications
with concurrency-sensitive operations, requiring unavailable,
synchronous coordination.

@@

\item For correct behavior, applications may require a combination of
  HAT and (ideally sparing use of) non-HAT isolation levels; future
  database designers should plan accordingly.
\end{introenumerate}

### 2013-10-07 writeup/model.tex
Final edits

In addition to high availability, which allows operations on any
replica, distributed algorithms often assume a model in which clients
always contact the same logical replica(s) across subsequent
operations, whereby each of the client's prior operations (but not
necessarily other clients' operations) are reflected in the database
state that they observe. As we will discuss in Section~\ref{sec:hats},
clients can ensure continuity between operations (e.g., reading their
prior updates to a data item) by maintaining affinity or
``stickiness'' with a server or set of servers~\cite{vogels-defs}. In
a fully replicated system, where all servers are replicas for all data
items, stickiness is simple: a client can maintain stickiness by
contacting the same server for each of its requests. However, to stay
``sticky'' in a partially-replicated system, where servers are
replicas for subsets of the set of data items (which we consider in
this paper), a client must maintain stickiness with a single
\textit{logical} copy of the database, which may consist of multiple
physical servers. We say that a system provides \textbf{sticky
  availability} if, whenever a client's transactions is executed
against a copy of database state that reflects all of the client's prior
transactions, it eventually receives a response, even in the presence
of indefinitely long partitions. A client may choose to become sticky
available by acting as a server itself; for example, a client might
cache its reads and writes~\cite{bolton, sessionguarantees,
  swift}. Any guarantee achievable in a highly available system is
achievable in a sticky high availability system but not vice-versa.

>>>>

In addition to high availability, which allows operations on any
replica, distributed algorithms often assume a model in which clients
always contact the same logical replica(s) across subsequent
operations, whereby each of the client's prior operations (but not
necessarily other clients' operations) are reflected in the database
state that they observe. As we will discuss in Section~\ref{sec:hats},
clients can ensure continuity between operations (e.g., reading their
prior updates to a data item) by maintaining affinity or
``stickiness'' with a server or set of servers~\cite{vogels-defs}. In
a fully replicated system, where all servers are replicas for all data
items, stickiness is simple: a client can maintain stickiness by
contacting the same server for each of its requests. However, to stay
``sticky'' in a partially-replicated system, where servers are
replicas for subsets of the set of data items (which we consider in
this paper), a client must maintain stickiness with a single
\textit{logical} copy of the database, which may consist of multiple
physical servers. We say that a system provides \textbf{sticky
  availability} if, whenever a client's transactions is executed
against a copy of database state that reflects all of the client's
prior operations, it eventually receives a response, even in the
presence of indefinitely long partitions (where ``reflects'' is
dependent on semantics). A client may choose to become sticky
available by acting as a server itself; for example, a client might
cache its reads and writes~\cite{bolton, sessionguarantees,
  swift}. Any guarantee achievable in a highly available system is
achievable in a sticky high availability system but not vice-versa.

### 2013-10-07 writeup/motivation.tex
Final edits

Why does high availability matter? Peter Deutsch starts his classic
list of ``Fallacies of Distributed Computing'' with two concerns
fundamental to distributed database systems: ``\textit{1.)}  The
network is reliable. \textit{2.)} Latency is
zero''~\cite{fallacies-deutsch}; in a distributed setting, network
failures may prevent database servers from communicating, and, in the
absence of failures, communication is slowed by factors like physical
distance, network congestion, and routing. As we will see
(Section~\ref{sec:availability}), highly available system design
mitigates the effects of network \textit{partitions} and
\textit{latency}. In this section, we draw on a range of evidence that
indicates that partitions occur with frequency in real-world
deployments and latencies between datacenters are substantial, often
on the order of several hundreds of milliseconds.

@@

Several recent studies rigorously quantify partition behavior. A 2011
study of several Microsoft datacenters observed over 13,300 network
failures with end-user impact, with an estimated median 59,000 packets
lost per failure. The study found a mean of 40.8 network link failures
per day (95th percentile: 136), with a median time to repair of around
five minutes (and up to one week). Perhaps surprisingly, provisioning
redundant networks only reduces impact of failures by up to 40\%,
meaning network providers cannot easily curtail partition
behavior~\cite{sigcomm-dc}. A 2010 study of over 200 wide-area routers
found an average of 16.2--302.0 failures per link per year with an
average annual downtime of 24--497 minutes per link per year (95th
percentile at least 34 hours)~\cite{sigcomm-wan}. In HP's managed
enterprise networks, WAN, LAN, and connectivity problems account for
28.1\% of all customer support tickets while 39\% of tickets relate to
network hardware.  The median incident duration for highest priority
tickets ranges from 114--188 minutes and up to a full day for all
tickets~\cite{turner2012failure}. Other studies confirm these results,
showing median time between connectivity failures over a WAN network
of approximately 3000 seconds with a median time to repair between 2
and 1000 seconds~\cite{ip-backbone-failures} as well as frequent path
routing failures on the Internet~\cite{labovitz-failures}. Finally,
Kingsbury and Bailis have recently collected an informal survey of
network failures~\cite{aphyr-post}. Isolating, quantifying, and
accounting for these network failures is an area of active research in
networking community~\cite{uw-failure-networks}.

@@

In real deployments, messages travel slower than the speed of light
due to routing, congestion, and server-side overheads. To illustrate
the difference between intra-datacenter, inter-datacenter, and
inter-planetary networks, we performed a measurement study of network
behavior on Amazon's EC2, a widely used public compute cloud. We
measured one week of ping times (i.e., round-trip times, or RTTs)
between all seven EC2 geographic ``regions,'' across three
``availability zones'' (closely co-located datacenters), and within a
single ``availability zone'' (datacenter), at a granularity of
1s\footnote{Data available at
  \url{http://bailis.org/blog/communication-costs-in-real-world-networks/}}. We
summarize the results of our network measurement study in
Table~\ref{table:rtt}. On average, intra-datacenter communication
(Table 1a) is between 1.82 and 6.38 times faster than across
geographically co-located datacenters (Table 1b) and between 40 and
647 times faster than across geographically distributed datacenters
(Table 1c). The cost of wide-area communication exceeds the speed of
light: for example, while a speed-of-light RTT from S\~{a}o Paulo to
Singapore RTT is 106.7ms, ping packets incur an average 362.8ms RTT
(95th percentile: 649ms). As shown in Figure~\ref{fig:rtt}, the
distribution of latencies varies between links, but the trend is
clear: remote communication has a substantial cost. As with network
failures, quantifying and minimizing communication delays is an active
area of research in the networking community~\cite{bobtail}.

@@

The previous section demonstrated that distributed systems must
address partitions and latency: what does this mean for distributed
databases? Database researchers and designers have long realized that
serializability is not achievable in a highly available
system~\cite{davidson-survey}, meaning that, in environments like
those in Section~\ref{sec:motivation}, database designs face a choice
between availability and strong semantics. However, even in a
single-node database, the coordination penalties associated with
serializability can be severe, manifesting themselves in the form of
decreased concurrency (and, subsequently, performance degradation,
scalability limitations, and, often, external aborts like
deadlocks)~\cite{gray-isolation}. Accordingly, to increase
concurrency, database systems offer a range of ACID properties weaker
than serializability: the host of so-called \textit{weak isolation}
models describe varying restrictions on the space of schedules that
are allowable by the system~\cite{adya, ansi-sql, ansicritique}. None
of these weak isolation models guarantees serializability, but, as we see below,
their benefits are often considered to outweigh costs of possible
consistency anomalies that might arise from their use.

>>>>

Why does high availability matter? Peter Deutsch starts his classic
list of ``Fallacies of Distributed Computing'' with two concerns
fundamental to distributed database systems: ``\textit{1.)}  The
network is reliable. \textit{2.)} Latency is
zero''~\cite{fallacies-deutsch}. In a distributed setting, network
failures may prevent database servers from communicating, and, in the
absence of failures, communication is slowed by factors like physical
distance, network congestion, and routing. As we will see
(Section~\ref{sec:availability}), highly available system designs
mitigate the effects of network \textit{partitions} and
\textit{latency}. In this section, we draw on a range of evidence that
indicates that partitions occur with frequency in real-world
deployments and latencies between datacenters are substantial, often
on the order of several hundreds of milliseconds.

@@

Several recent studies rigorously quantify partition behavior. A 2011
study of several Microsoft datacenters observed over 13,300 network
failures with end-user impact, with an estimated median 59,000 packets
lost per failure. The study found a mean of 40.8 network link failures
per day (95th percentile: 136), with a median time to repair of around
five minutes (and up to one week). Perhaps surprisingly, provisioning
redundant networks only reduces impact of failures by up to 40\%,
meaning network providers cannot easily curtail partition
behavior~\cite{sigcomm-dc}. A 2010 study of over 200 wide-area routers
found an average of 16.2--302.0 failures per link per year with an
average annual downtime of 24--497 minutes per link per year (95th
percentile at least 34 hours)~\cite{sigcomm-wan}. In HP's managed
enterprise networks, WAN, LAN, and connectivity problems account for
28.1\% of all customer support tickets while 39\% of tickets relate to
network hardware.  The median incident duration for highest priority
tickets ranges from 114--188 minutes and up to a full day for all
tickets~\cite{turner2012failure}. Other studies confirm these results,
showing median time between connectivity failures over a WAN network
of approximately 3000 seconds with a median time to repair between 2
and 1000 seconds~\cite{ip-backbone-failures} as well as frequent path
routing failures on the Internet~\cite{labovitz-failures}. A recent,
informal report by Kingsbury and Bailis catalogs a host of additional
practitioner reports~\cite{aphyr-post}. Not surprisingly, isolating,
quantifying, and accounting for these network failures is an area of
active research in networking community~\cite{uw-failure-networks}.

@@

In real deployments, messages travel slower than the speed of light
due to routing, congestion, and server-side overheads. To illustrate
the difference between intra-datacenter, inter-datacenter, and
inter-planetary networks, we performed a measurement study of network
behavior on Amazon's EC2, a widely used public compute cloud. We
measured one week of ping times (i.e., round-trip times, or RTTs)
between all seven EC2 geographic ``regions,'' across three
``availability zones'' (closely co-located datacenters), and within a
single ``availability zone'' (datacenter), at a granularity of
1s\footnote{Data available at {\scriptsize
    \url{http://bailis.org/blog/communication-costs-in-real-world-networks/},
    \url{https://github.com/pbailis/aws-ping-traces}}}. We summarize
the results of our network measurement study in
Table~\ref{table:rtt}. On average, intra-datacenter communication
(Table 1a) is between 1.82 and 6.38 times faster than across
geographically co-located datacenters (Table 1b) and between 40 and
647 times faster than across geographically distributed datacenters
(Table 1c). The cost of wide-area communication exceeds the speed of
light: for example, while a speed-of-light RTT from S\~{a}o Paulo to
Singapore RTT is 106.7ms, ping packets incur an average 362.8ms RTT
(95th percentile: 649ms). As shown in Figure~\ref{fig:rtt}, the
distribution of latencies varies between links, but the trend is
clear: remote communication has a substantial cost. Quantifying and
minimizing communication delays is also an active area of research in
the networking community~\cite{bobtail}.

@@

The previous section demonstrated that distributed systems must
address partitions and latency: what does this mean for distributed
databases? Database researchers and designers have long realized that
serializability is not achievable in a highly available
system~\cite{davidson-survey}, meaning that, in environments like
those in Section~\ref{sec:motivation}, database designs face a choice
between availability and strong semantics. However, even in a
single-node database, the coordination penalties associated with
serializability can be severe and are manifested in the form of
decreased concurrency (and, subsequently, performance degradation,
scalability limitations, and, often, aborts due to deadlock or
contention)~\cite{gray-isolation}. Accordingly, to increase
concurrency, database systems offer a range of ACID properties weaker
than serializability: the host of so-called \textit{weak isolation}
models describe varying restrictions on the space of schedules that
are allowable by the system~\cite{adya, ansi-sql, ansicritique}. None
of these weak isolation models guarantees serializability, but, as we
see below, their benefits are often considered to outweigh costs of
possible consistency anomalies that might arise from their use.
